---
title: "Laboratory Exercise Week 1"
author: "Brian Tipton"
date: "8/23/23"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*:

-   Write your R code inside the code chunks after each question.
-   Write your answer comments after the `#` sign.
-   To generate the word document output, click the button `Knit` and wait for the word document to appear.
-   RStudio will prompt you (only once) to install the `knitr` package.
-   Submit your completed laboratory exercise using Blackboard's Turnitin feature. Your Turnitin upload link is found on your Blackboard Course shell under the Laboratory folder.

------------------------------------------------------------------------

1.  Create a vector of three elements `(2,4,6)` and name that vector `vec.a`. Create a second vector, `vec.b`, that contains `(8,10,12)`.
    a.  Add these two vectors together and name the result `vec.c`.
    b.  Create a vector, named `vec.d`, that contains only two elements `(14,20)`. Add this vector to `vec.a`. What is the result and what do you think `R` did (look up the recycling rule using `Google`)? What is the warning message that `R` gives you?
    c.  Next add 5 to the vector `vec.a`. What is the result and what did `R` do? Why doesn't it give you a warning message similar to what you saw in the previous problem?

*My Custom functions used from my local lab projects .Rprofile*

```{r message=FALSE}
source("../.Rprofile", chdir = TRUE)
```

```{r}
catXWithString
```

## Code chunk

```{r message=TRUE}
# Insert your code for this question after this line
trial.vec <- 1:20
# trial.vec
vec.a <- c(2, 4, 6)
vec.b <- c(8, 10, 12)
vec.c <- vec.a + vec.b
vec.d <- c(14, 20)

(vec.a + vec.d) |>
  catXWithString(string = "(VEC.a + VEC.d) -> ")

catXWithString("VEC.c: ", vec.c)

(vec.a + 5) |>
  catXWithString(string = "(VEC.a + 5) -> ")
```

2.  Generate the vector of even numbers `{2, 4, 6, . . . , 20}`
    a.  Using the `seq()` function and
    b.  Using the `a:b` shortcut and some subsequent algebra. *Hint: Generate the vector 1-10 and then multiple it by 2.*

```{r}
seq(from = 2, to = 20, by = 2) |>
  catXWithString(string = "ONE TO TWENTY EVENS USING SEQ:\n\t")

2:20 |>
  (\(x) x[x %% 2 == 0])() |>
  catXWithString(string = "ONE TO TWENTY EVENS USING MODULO:\n\t")
```

3.  Create a vector `y` containing `(2, 2, 2, 2, 4, 4, 4, 4, 8, 8, 8, 8)` using the `rep()` function. You might need to check the help file for `rep()` by typing `?rep` in the console to see all of the options that `rep()` will accept. In particular, look at the optional argument `each=`.
    a.  Find the mean of vector `y` using the function `mean()`.
    b.  Use google search to find the function in `R` that computes the variance of a vector and find the variance of `y`.

```{r}
y <- rep(2^(1:3), each = 4)

catXWithString("Y: ", y)

mean(y) |>
  catXWithString(string = "MEAN OF Y: ")
var(y) |>
  catXWithString(string = "VARIANCE OF Y: ")
```

4.  The vector `letters` is a built-in vector to `R` and contains the lower case English alphabet.
    a.  Extract the 9th element of the `letters` vector.
    b.  Extract the sub-vector that contains the 9th, 11th, and 19th elements.
    c.  Extract the sub-vector that contains everything except the last two elements.

```{r}
letters[9] |>
  catXWithString(string = "NINETH LETTER ALPHA: ")

letters[c(9, 11, 19)] |>
  catXWithString(
    string =
      "Alpha Subvector 9th, 11th and 19th indice: "
  )

head(letters, -2) |>
  catXWithString(string = "EVERYTHING BUT LAST 2: ")
```
