---
title: "Laboratory Exercise Week 2"
author: "Brian Tipton"
date: "8/30/23"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Directions*: 

* Write your R code inside the code chunks after each question.
* Write your answer comments after the `#` sign.
* To generate the word document output, click the button `Knit` and wait for the word document to appear.
* RStudio will prompt you (only once) to install the `knitr` package.
* Submit your completed laboratory exercise using Blackboard's Turnitin feature. Your Turnitin upload link is found on your Blackboard Course shell under the Laboratory folder.

*My Custom functions used from my local lab projects .Rprofile*

```{r message=FALSE}
source("../../.Rprofile", chdir = TRUE)
```

```{r}
catXWithString
```
***

1. You will work with a matrix whose entries are all prime numbers below 30.  
      i)  Create this matrix using the function `matrix()` with five rows. Save the matrix as `P`.  
      ii) Extract the second and third row out of `P`.  
      iii) Extract the entry in the fourth row and first column of `P`.  
      iv) Generate the transpose of the matrix using the function `t()`. What is the new dimension of this matrix.  

### Code chunk
```{r}
P <- matrix(c(2, 3, 5, 7, 11, 13, 17, 19, 23, 29), nrow = 5)
P
cat("SECOND AND THIRD ROWS: \n")
P[c(2,3),]
cat("\n")
P[4,1] |> catXWithString(string = "ITEM AT ROW 4 COLUMN 1:")
tP <- t(P)
cat("Transposed Matrix: \n")
tP
cat("\n")
catXWithString(nrow(tP), string = "NUM ROWS:")
catXWithString(ncol(tP), string = "NUM COL:")
```

2. Create and manipulate a data frame.
      i) Create a `data.frame` named `my.trees` that has the following columns:  
          - `Girth = c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0)`  
          - `Height = c(70, 65, 63, 72, 81, 83, 66)`  
          - `Volume = c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)`  
      ii) Extract the fifth observational unit.  
      iii) Extract the `Girth` column referring to it by name.  
      iv) Print out a data frame of all the observations except for the last observation.  


### Code chunk
```{r}
my.trees <-
  data.frame(
    Girth = c(8.3, 8.6, 8.8, 10.5, 10.7, 10.8, 11.0),
    Height = c(70, 65, 63, 72, 81, 83, 66),
    Volume = c(10.3, 10.3, 10.2, 16.4, 18.8, 19.7, 15.6)
  )
cat("5th Observational Unit")
my.trees[5,]
cat("\n")
cat("GIRTH COLUMN: ")
my.trees["Girth"] 
cat("\n")
cat("ALL EXCEPT LAST OBSERVATION: ")
my.trees[1:(nrow(my.trees) - 1),]
cat("\n")
```

3. The popular `iris` data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica. The `iris` data set is included with every R installation.  
      i) Check the structure of the`iris` data.   
      ii) How many variables and observations are in the data set?   
      iii) Which variables are `numeric` type?  
      iv) Display the first 4 rows of the data.  



### Code chunk
```{r} 
ncol(iris) |> catXWithString(string = "How many variables: ")
nrow(iris) |> catXWithString(string = "How many observations: ")
is.numeric(iris$Sepal.Length) |> catXWithString(string = "Sepal.Length is Numeric: ")
is.numeric(iris$Sepal.Width) |> catXWithString(string = "Sepal.Width is Numeric: ")
is.numeric(iris$Petal.Length) |> catXWithString(string = "Petal.Length is Numeric: ")
is.numeric(iris$Petal.Width) |> catXWithString(string = "Petal.Width is Numeric: ")
is.numeric(iris$Species) |> catXWithString(string = "Species is Numeric: ")
cat("FIRST 4 ROWS: ")
iris[1:4,]
```



  
4.  You will work with the passenger Titanic data found on this [link](<https://goo.gl/NHb1Pg>).   
      i) Read this data directly from the web using its URL.  
      ii) Read this data by saving it first into your working directory. Check your working directory using `getwd()` or change its location using the steps covered in the lesson.  
      iii) How many passengers and what variables are in the data set?   
      iv) Display the first 5 rows of the data.  

### Code chunk
```{r}
titanic.remote <- read.csv("https://goo.gl/NHb1Pg")
titanic.local <- read.csv("./titanic.csv")
nrow(titanic.remote) |> catXWithString(string = "NUM PASSENGERS: ")
names(titanic.remote) |> catXWithString(string = "VARIABLES: ")
titanic.remote[1:5,]
```

